Rapid-response Bayesian sky localization for electromagnetic follow-up of gravitational-wave candidates

### 2012-07-31 ms.tex
First update of ms.tex

>>>>

\section{New section, add a new name to this}

A CBC will emit a GW. What we detect will be the data, $x(t)$ which is a function of time. This data will be a combination of the signal emitted from the CBC< $h(t)$, and the noise our detector picks up in addition to the signal, $n(t)$. We assume we have additive, zero mean, wide source stationary, Gaussian measurement noise $n(t)$. This means that noise is only added to what passes through the channel, the mean of this noise is zero, the noise does not vary in time, and the noise is drawn from a Gaussian distribution.

Our goal here is to first determine the likelihood.

Our detector data is given by $x(t)$. This can be described in the frequency domain as $\tilde{x}(f) = \tilde{h}(f) + \tilde{n}(f)$ where $\tilde{h}(f)$ is our signal and $\tilde{n}(f)$ is the noise. We have noise, $E[n(t_1)n(t_2)]$ which depends only on $t_1 - t_2$. The Fourier transform of the noise gives us this: $E[\tilde{n}(f)^*\tilde{n}(f)] = \delta (f_1 - f_2) s(f_1)$ where $s(f)$ is the power spectral density for the noise. The power spectral density gives us information about the energy of the present frequencies. The delta function is only non-zero when the input is zero, so in this case we will only get a nonzero value for $\delta (f_1 - f_2) s(f_1)$ when $f_1 = f_2$.

To determine the power spectral density (PSD) for the real signal $y(t)$ we first take the Fourier transform of our real signal to get:

\begin{equation}
	\tilde{y}(f) = \int_{-\infty}^{\infty}y(t)e^{-2\pi ift}dt
\end{equation}

Then we get the two sided PSD from the definition of PSD.

\begin{equation}
	R_y(f) = |\tilde{y}(f)|^2
\end{equation}

Then we get the single sided PSD with a conversion,

\begin{equation}
	S_y(f) = |\tilde{y}(f)|^2 + |\tilde{y}(-f)|^2
\end{equation}

We can decompose our signal into two orthogonal bases with

\begin{equation}
	\tilde{h}(f) = a\tilde{h}_0(f) + b\tilde{h}_{\pi/2}(f)
\end{equation}

Defining
\begin{eqnarray}
	\tilde{h}_0(f) & = & -i \tilde{h}_{\pi/2}(f) \mbox{ for } f > 0 \\
	\tilde{h}_0(-f) & = & i \tilde{h}_{\pi/2}(-f) \mbox{ for } f < 0
\end{eqnarray}

This ends up resulting in
\begin{equation}
	\tilde{h}(f) = (a + bi)\tilde{h}_0(f)
\end{equation}

Next we'll set $\tilde{u}(f) = \tilde{h}_0(f)$.

We can rewrite our signal in polar form to get the following.
\begin{eqnarray}
	\tilde{h}(f) & = & (a + bi) \tilde{u}(f) e^{-2\pi ift_0} \\
	& = & Ce^{i \phi}\tilde{u}(f)e^{-2\pi ift_0} \\
	& = & Ce^{i(\phi - 2\pi ft_0)}\tilde{u}(f)
\end{eqnarray}

This gives us a definition of $\tilde{h}(f)$ to be used in our likelihood and allows us to update our definition of $\tilde{x}(f)$

\begin{eqnarray}
	\tilde{x}(f) & = & Ce^{i(\phi - 2\pi ft_0)}\tilde{u}(f) + \tilde{n}(f) \mbox{ for } f > 0 \\
	\tilde{x}(f) & = & \tilde{x}^*(f) \mbox{ for } f < 0
\end{eqnarray}

We can now obtain our likelihood function from the definition of a univariate gaussian for a continuous signal ${\bf{x}}(f)$,
\begin{equation}
	f({\bf{x}}) = \frac{1}{\sqrt{\det(2\pi \bf{\Sigma})}}\exp(-\frac{1}{2}(\bf{x}-\bf{u})^T\Sigma^{-1}(\bf{x}-\bf{\mu}))
\end{equation}
Expressing this in frequency space,
\begin{equation}
	f({\bf{\tilde{x}}}(f)) \propto \exp[-\frac{1}{2}\int\int( {\bf{x}}(f)-{\bf{\mu}}(f))^*C(f,f')(\tilde{x}(f')-\mu(f'))]dfdf'
\end{equation}
Where $C(f,f')$ represents the covariance. Now we can determine the log of our likelihood function.

\begin{equation}
	\ln(L(\tilde{x}; C, \phi, t_0)) \sim -\frac{1}{2}\int_0 ^\infty \frac{|\tilde{x}(f) - \tilde{h}(f; C, \phi, t_0)|^2}{S_n(f)}df
\end{equation}

There will be a time delay between the time of coalescence and the time the signal arrives at our detector. We'll see what this means we should expect by introducing a time delay in frequency space. Our real signal data is given by $y(t)$, so we can introduce a delay by changing the time, $y(t - t_0) = z(t)$ where $t$ is the time of arrival and $t_0$ the time of coalescence. So $z(t)$ now contains our time delay. We take the Fourier transform of $z(t)$ and are left with

\begin{equation}
	\tilde{z}(f) = \int^{\infty}_{-\infty}y(t-t_0)e^{-2\pi ift} dt
\end{equation}

Introducing a change of variables from $t$ to $t' = t - t_0$ gives us
\begin{eqnarray}
	\tilde{z}(f) & = & e^{-2\pi ift_0} \int_{-\infty}^{\infty}y(t')e^{-2\pi ift'} dt' \\
	& = & e^{-2\pi ift_0} \tilde{y}(f)
\end{eqnarray}

This means that a time delay introduced a phase shift in frequency.

Now we can develop the Fisher information matrix. This will tell us how our variables affect each other. The Fisher information matrix is defined as 
\begin{equation}
\mathcal{I}_{ij} \equiv E[-\frac{\partial^2 \ln L}{\partial \theta_i \partial \theta_j}]
\end{equation}
 for $L(\theta_1, \theta_2, ...)$.

So if we take our likelihood we'll end up with this,
\begin{eqnarray}
	\frac{\partial \ln L}{\partial \theta_i} & = & -\frac{1}{2} \int_0 ^\infty \frac{2}{s_n(f)}Re\{[\frac{\partial}{\partial \theta_i} (\tilde{x} - \tilde{h})][\tilde{x} - \tilde{h}]\}df \\
	& = & \int_0 ^\infty Re[(\frac{\partial \tilde{h}}{\partial \theta_i})^*(\tilde{x} - \tilde{h})]\frac{df}{s(f)} \\
	\frac{\partial^2 \ln L}{\partial \theta_i \partial \theta_j} & = & \int_0 ^\infty Re[(\frac{\partial^2 \tilde{h}}{\partial \theta_i \partial \theta_j})^*(\tilde{x} - \tilde{h})\frac{df}{s(f)} + \int_0 ^\infty Re[(\frac{\partial \tilde{h}}{\partial \theta_i})^*\frac{\partial \tilde{h}}{\partial \theta_j}]\frac{df}{s(f)}] \\
\end{eqnarray}

Apply this to the definition of the Fisher information matrix,
\begin{eqnarray}
	\mathcal{I}_{ij} & = & E[-\int_0^\infty (\frac{\partial ^2 \tilde{h}}{\partial \theta_i \partial \theta_j})^*(\tilde{x} - \tilde{h})\frac{df}{s(f)} + \int_0 ^ \infty Re[(\frac{\partial \tilde{h}}{\partial \theta_i})^*\frac{\partial \tilde{h}}{\partial \theta _j}]\frac{df}{s(f)}] \\
	& = & \int^\infty _0 Re[(\frac{\partial \tilde{h}}{\partial \theta_i})^* \frac{\partial \tilde{h}}{\partial \theta _j}]\frac{df}{s(f)}
\end{eqnarray}

Putting this into a matrix and defining $M_k = \int_0^\infty \tilde{u}(f)^* u(f) f\frac{df}{s(f)}$ and $N_k = \frac{M_k}{M_0}$ gives us
\[
\mathcal{I} =
\left( {\begin{array}{ccc}
 M_0 & 0 & 0 \\
 0 & C^2M_0 & -2\pi C^2M_1 \\
 0 & -2\pi C^2M_1 & 4\pi^2C^2M_2
 \end{array} } \right)
\]

Now we will perform three changes of variables to diagonalize this matrix.

The first change of variables is from $C$ to $\rho = \sqrt{M_0}C$. 
\[
\mathcal{I} =
\left( {\begin{array}{ccc}
 1 & 0 & 0 \\
 0 & \rho^2 & -2\pi\rho^2 N_1 \\
 0 & -2\pi\rho^2 N_1 & 4\pi^2 \rho^2 N_2
 \end{array} } \right)
\]

The second from $t$ to $\phi_t = t2\pi \sqrt{N_2}$ gives us
\[
\mathcal{I} =
\left( {\begin{array}{ccc}
 1 & 0 & 0 \\
 0 & \rho^2 & -\rho^2\frac{N_1}{\sqrt{N_2}} \\
 0 & -\rho^2\frac{N_1}{\sqrt{N_2}} & \rho^2
 \end{array} } \right)
\]

The third introduces two new variables, $\phi_s = \frac{1}{\sqrt{2}}(\phi + \phi_t)$ and $\phi_d = \frac{1}{\sqrt{2}}(\phi - \phi_t)$. This results in this matrix,
\[
\mathcal{I} =
\left( {\begin{array}{ccc}
 1 & 0 & 0 \\
 0 & (1 - \frac{N_1}{\sqrt{N_2}})\rho^2 & 0 \\
 0 & 0 & (1 + \frac{N_1}{\sqrt{N_2}})\rho^2
 \end{array} } \right)
\]

Now we'll do a final change of variables to make the phases have the same variance. We let $\alpha_s = \phi_s \sqrt{1 - \frac{N_1}{\sqrt{N_2}}}$ and $\alpha_d = \phi_d \sqrt{1 - \frac{N_1}{\sqrt{N_2}}}$ which gives us the matrix
\[
\mathcal{I} =
\left( {\begin{array}{ccc}
 1 & 0 & 0 \\
 0 & \rho^2 & 0 \\
 0 & 0 & \rho^2
 \end{array} } \right)
\]
